### 2015-12-08 ada-ftrl-proofs.tex
Added proofs for mirror descent section.
Use small-caps for all algorithms names
Make AdaFTRL recurrence a separate lemma

\section{Proofs for AdaFTRL}
\label{section:ada-ftrl-proof}

@@

\begin{proof}[Proof of Lemma~\ref{lemma:recurrence-solution}]
Let $a_t = \|\ell_t\|_* \max\{D, 1/\sqrt{2\lambda}\}$. The statement of the
lemma is equivalent to $\Delta_T \le \sqrt{3 \sum_{t=1}^T a_t^2}$ which we
prove by induction on $T$.  The base case $T=0$ is trivial. For $T \ge 1$, we
have
$$
\Delta_T
\le \Delta_{T-1} + \min \left\{a_T, \ \frac{a_T^2}{\Delta_{T-1}} \right\}
\le \sqrt{3 \sum_{t=1}^{T-1} a_t^2} + \min \left\{ a_T, \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
$$
where the first inequality follows from Lemma~\ref{lemma:gap-recurrence}, and
the second inequality from the induction hypothesis and the fact that $f(x) = x
+ \min\{a_T, a_T^2/x\}$ is an increasing function of $x$.  It remains to prove
that
$$
\sqrt{3 \sum_{t=1}^{T-1} a_t^2} + \min \left\{ a_T, \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
\le  \sqrt{3 \sum_{t=1}^T a_t^2} \; .
$$
Dividing through by $a_T$ and making substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1} a_t^2}}{a_T}$, leads to
$$
z\sqrt{3} + \min\left\{1,\frac{1}{z\sqrt{3}}\right\} \le \sqrt{3 + 3z^2}
$$
which can be easily checked by considering separately the cases $z \in
[0,\frac{1}{\sqrt{3}})$ and $z \in [\frac{1}{\sqrt{3}}, \infty)$.
\end{proof}

>>>>

\section{Proofs for \textsc{AdaFTRL}}
\label{section:ada-ftrl-proofs}

@@

\begin{proof}[Proof of Lemma~\ref{lemma:recurrence-solution}]
We prove the lemma by induction on $T$.  The base case $T=0$ is trivial. For $T
\ge 1$, we have
$$
\Delta_T
\le \Delta_{T-1} + \min \left\{a_T, \ \frac{a_T^2}{\Delta_{T-1}} \right\}
\le \sqrt{3 \sum_{t=1}^{T-1} a_t^2} + \min \left\{ a_T, \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
$$
where the first inequality follows from Lemma~\ref{lemma:gap-recurrence}, and
the second inequality from the induction hypothesis and the fact that $f(x) = x
+ \min\{a_T, a_T^2/x\}$ is an increasing function of $x$.  It remains to prove
that
$$
\sqrt{3 \sum_{t=1}^{T-1} a_t^2} + \min \left\{ a_T, \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
\le  \sqrt{3 \sum_{t=1}^T a_t^2} \; .
$$
Dividing through by $a_T$ and making substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1} a_t^2}}{a_T}$, leads to
$$
z\sqrt{3} + \min\left\{1,\frac{1}{z\sqrt{3}}\right\} \le \sqrt{3 + 3z^2}
$$
which can be easily checked by considering separately the cases $z \in
[0,\frac{1}{\sqrt{3}})$ and $z \in [\frac{1}{\sqrt{3}}, \infty)$.
\end{proof}

### 2015-12-08 ada-ftrl.tex
Make AdaFTRL recurrence a separate lemma

\begin{lemma}[Initial Regret Bound]
\label{lemma:initial-regret-bound}
\textsc{AdaFTRL}, for any $u \in K$ and any $u \ge 0$, satisfies
$\Regret_T(u) \le \left(1 + R(u) \right) \Delta_T$.
\end{lemma}

@@

\begin{proof}
The inequality results from strong convexity of $R_t(w)$ and
Proposition~\ref{proposition:conjugate-properties}.
\end{proof}

\begin{lemma}[Solution of the Recurrence]
\label{lemma:recurrence-solution}
Let $D$ be the diameter of $K$. The sequence $\{\Delta_t\}_{t=0}^\infty$
generated by \textsc{AdaFTRL} satisfies for any $T \ge 0$,
$$
\Delta_T \le \sqrt{3} \max\left\{D, \frac{1}{\sqrt{2\lambda}} \right\} \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
\end{lemma}
%
Proof of the Lemma~\ref{lemma:recurrence-solution} is deferred
to~\ref{section:ada-ftrl-proofs}.  Theorem~\ref{theorem:ada-ftrl-regret-bound}
follows from
Lemmas~\ref{lemma:initial-regret-bound}~and~\ref{lemma:recurrence-solution}.

>>>>

\begin{lemma}[Initial Regret Bound]
\label{lemma:initial-regret-bound}
\textsc{AdaFTRL} satisfies, for any $u \in K$ and any $T \ge 0$,
$$
\Regret_T(u) \le \left(1 + R(u) \right) \Delta_T \; .
$$
\end{lemma}

@@

\begin{proof}
By definition, $\Delta_t$ satisfies the recurrence
$$
\Delta_t = \Delta_{t-1} + \Delta_{t-1} \Breg_{R^*}\left( - \frac{L_t}{\Delta_{t-1}}, - \frac{L_{t-1}}{\Delta_{t-1}} \right) \; .
$$
Using parts 4 and 6 of Proposition~\ref{proposition:conjugate-properties}, we
can upper bound $\Breg_{R^*}\left( - \frac{L_t}{\Delta_{t-1}}, -
\frac{L_{t-1}}{\Delta_{t-1}} \right)$ with two different quantities.  Taking
minimum of the two quantities finishes the proof.
\end{proof}

The recurrence of Lemma~\ref{lemma:gap-recurrence} can be simplified. Defining
$$
a_t = \|\ell_t\|_* \max \left\{D, \frac{1}{\sqrt{2 \lambda}} \right\} \; ,
$$
we get a recurrence
$$
\Delta_t \le \Delta_{t-1} + \min \left\{ a_t, \ \frac{a_t^2}{\Delta_{t-1}} \right\} \; .
$$
The next lemma solves this recurrence, by giving an explicit upper bound on
$\Delta_T$ in terms of $a_1, a_2, \dots, a_T$.

\begin{lemma}[Recurrence]
\label{lemma:recurrence-solution}
Let $\{a_t\}_{t=1}^\infty$ be any sequence of non-negative real numbers.
Suppose that $\{\Delta_t\}_{t=0}^\infty$ is a sequence of non-negative real
numbers satisfying
$$
\Delta_0 = 0 \qquad \text{and} \qquad
\Delta_t \le \Delta_{t-1} + \min \left\{ a_t, \ \frac{a_t^2}{\Delta_{t-1}} \right\} \quad \text{for any $t \ge 1$} \; .
$$
Then, for any $T \ge 0$,
$$
\Delta_T \le \sqrt{3 \sum_{t=1}^T a_t^2} \; .
$$
\end{lemma}
%
Proof of the Lemma~\ref{lemma:recurrence-solution} is deferred
to~\ref{section:ada-ftrl-proofs}.  Theorem~\ref{theorem:ada-ftrl-regret-bound}
follows from
Lemmas~\ref{lemma:initial-regret-bound},~\ref{lemma:gap-recurrence}~and~\ref{lemma:recurrence-solution}.
